---
title: "WhatsApp Chat Sentiment Analysis"
author: "Alexander Oswald"
date: "2025-03-24"
output: ioslides_presentation
---

```{r setup, include=FALSE}
# Load required libraries
library(tidyverse)
library(lubridate)

# Read the chat file
file_path <- paste0(Sys.getenv("WHATSAPP_TEXTS_DIR"),"_chat.txt")  # Change to the correct file path
lines <- readLines(file_path, encoding = "UTF-8")

# Extract timestamp, sender, and message using regex
chat_df <- tibble(
  raw_text = lines
) %>%
  mutate(
    Date = str_extract(raw_text, "\\d{4}-\\d{2}-\\d{2}"),
    Time = str_extract(raw_text, "\\d{1,2}:\\d{2}:\\d{2} [APM]+"),
    Sender = str_extract(raw_text, "(?<=\\])\\s*[^:]+"),
    Message = str_extract(raw_text, "(?<=: ).+")
  ) %>%
  drop_na(Date, Sender, Message) %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d"))

# View first few rows
head(chat_df)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r Perform Sentiment Analysis, echo = TRUE}
library(syuzhet)

# Get sentiment scores
chat_df <- chat_df %>%
  mutate(
    Sentiment_Score = get_sentiment(Message, method="bing"),
    Sentiment = case_when(
      Sentiment_Score > 0 ~ "Positive",
      Sentiment_Score < 0 ~ "Negative",
      TRUE ~ "Neutral"
    )
  )

# View sentiment distribution
table(chat_df$Sentiment)

# Plot sentiment trends over time
chat_df %>%
  group_by(Date, Sentiment) %>%
  summarise(Count = n()) %>%
  ggplot(aes(x = Date, y = Count, color = Sentiment)) +
  geom_line() +
  labs(title = "Sentiment Trend Over Time", x = "Date", y = "Count of Messages") +
  theme_minimal()
```

## Detecting Abusive Messages

```{r Detect Abusive Messages}
# Define abusive keywords
abusive_keywords <- c("shut up", "idiot", "moron", "lazy", "selfish", "disrespect", 
                      "you never", "you always", "it's your fault", "stop texting me", 
                      "fk", "fuck", "annoying", "do not ever", "you're atrocious", 
                      "you're speaking with emotions", "are you a kid", "prove it")

# Flag abusive messages
chat_df <- chat_df %>%
  mutate(Abusive = if_else(str_detect(tolower(Message), paste(abusive_keywords, collapse="|")), TRUE, FALSE))

# View abusive messages
abusive_messages <- chat_df %>% filter(Abusive == TRUE)
print(abusive_messages)

# Plot abusive message trends over time
abusive_trend <- chat_df %>%
  group_by(Date) %>%
  summarise(Abusive_Count = sum(Abusive))

ggplot(abusive_trend, aes(x = Date, y = Abusive_Count)) +
  geom_line(color="red") +
  labs(title = "Abusive Messages Over Time", x = "Date", y = "Count of Abusive Messages") +
  theme_minimal()

```

## Check Word Cloud for Context
```{r Word Cloud Creation}
library(tidytext)
library(ggwordcloud)
library(gridExtra)

# Collapse messages by sender into a single text
messages_by_sender <- df %>%
  group_by(Sender) %>%
  summarise(all_text = paste(Message, collapse = " "), .groups = "drop")

# Tokenize and count words for each sender
word_counts <- messages_by_sender %>%
  unnest_tokens(word, all_text) %>%
  count(Sender, word, sort = TRUE)

# Split into separate data frames per sender
sender_list <- unique(word_counts$Sender)
wordclouds <- list()

for (sender in sender_list) {
  wordcloud_plot <- word_counts %>%
    filter(Sender == sender) %>%
    ggplot(aes(label = word, size = n)) +
    geom_text_wordcloud_area() +
    theme_minimal() +
    ggtitle(paste("Word Cloud for", sender))

  wordclouds[[sender]] <- wordcloud_plot
}

# Display side-by-side (adjust to number of senders)
do.call(grid.arrange, c(wordclouds, ncol = length(wordclouds)))
```

```{r Identify Negative Responsives to abusive messages}
# Identify cases where Alex responds negatively to an abusive message from Sun Shuo Pershing
reaction_patterns <- df %>%
  filter(
    Sender == "Alex",
    Sentiment == "Negative",
    Previous_Sender == "Sun Shuo Pershing",
    Previous_Abusive == TRUE
  )

# Display reaction-based patterns
print(reaction_patterns)
```

```{r Identify Negative Responsives to abusive messages}
# Identify cases where Alex responds negatively to an abusive message from Sun Shuo Pershing
reaction_patterns <- df %>%
  filter(
    Sender == "Alex",
    Sentiment == "Negative",
    Previous_Sender == "Sun Shuo Pershing",
    Previous_Abusive == TRUE
  )

# Display reaction-based patterns
print(reaction_patterns)
```

## Define Categorization
```{r Categorizations}
library(tidyverse)
library(stringr)

# Define keyword lists
abusive_keywords     <- c("abuse", "idiot", "stupid", "hate", "shut up")  # Define this ahead of time
positive_keywords    <- c("love", "thanks", "appreciate", "good job", "well done", "proud", "happy", "yay", "â¤ï¸")
neutral_keywords     <- c("hi", "hello", "how are you", "what's up", "okay", "good", "ðŸ‘ðŸ¼", "nice", "cool", "haha", "lol")
task_keywords        <- c("pick up", "bring", "buy", "appointment", "meeting", "schedule", "cook", "walk", "clean", "fix")
conflict_keywords    <- c("no", "disagree", "not true", "wrong", "but", "i don't think", "you didn't", "why")
request_keywords     <- c("please", "can you", "should", "need to", "have to", "must", "do it", "go get", "bring me")
apology_keywords     <- c("sorry", "apologize", "my bad", "forgive", "i messed up", "i was wrong")
info_keywords        <- c("sent", "emailed", "update", "news", "info", "details", "link", "check", "read")
humor_keywords       <- c("joking", "lol", "haha", "funny", "meme", "sarcastic", "lmao")

# Function to categorize a message
categorize_message <- function(text) {
  text <- tolower(text)

  if (any(str_detect(text, fixed(abusive_keywords)))) {
    return("Abusive")
  } else if (any(str_detect(text, fixed(positive_keywords)))) {
    return("Supportive/Positive")
  } else if (any(str_detect(text, fixed(neutral_keywords)))) {
    return("Neutral/Chit-chat")
  } else if (any(str_detect(text, fixed(task_keywords)))) {
    return("Task-Oriented")
  } else if (any(str_detect(text, fixed(conflict_keywords)))) {
    return("Disagreements/Conflict")
  } else if (any(str_detect(text, fixed(request_keywords)))) {
    return("Requests/Commands")
  } else if (any(str_detect(text, fixed(apology_keywords)))) {
    return("Apologies/Repair Attempts")
  } else if (any(str_detect(text, fixed(info_keywords)))) {
    return("Information Sharing")
  } else if (any(str_detect(text, fixed(humor_keywords)))) {
    return("Humor/Sarcasm")
  } else {
    return("Neutral/Chit-chat")
  }
}

# Apply the categorization function to the dataframe
df <- df %>%
  mutate(Category = map_chr(Message, categorize_message))

# View categorized messages (in RStudio)
df %>% view("Categorized Messages")

# Plot category distribution
df %>%
  count(Category) %>%
  ggplot(aes(x = fct_reorder(Category, n), y = n)) +
  geom_col(fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Distribution of Message Categories",
    x = "Category",
    y = "Count of Messages"
  ) +
  theme_minimal()

# Remove "Neutral/Chit-chat" column if it exists
filtered_category_by_sender <- category_by_sender %>%
  select(-`Neutral/Chit-chat`)

# View filtered data
filtered_category_by_sender %>% view("Filtered Category Distribution by Sender")

# Convert to long format for stacked bar plot
long_filtered <- filtered_category_by_sender %>%
  rownames_to_column("Sender") %>%  # If Sender was in row names
  pivot_longer(-Sender, names_to = "Category", values_to = "Count")

# Plot
ggplot(long_filtered, aes(x = Sender, y = Count, fill = Category)) +
  geom_col(position = "stack") +
  labs(
    title = "Message Category Distribution by Sender (Excluding Neutral/Chit-chat)",
    x = "Sender",
    y = "Count of Messages",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )# Remove "Neutral/Chit-chat" column if it exists
filtered_category_by_sender <- category_by_sender %>%
  select(-`Neutral/Chit-chat`)

# View filtered data
filtered_category_by_sender %>% view("Filtered Category Distribution by Sender")

# Convert to long format for stacked bar plot
long_filtered <- filtered_category_by_sender %>%
  rownames_to_column("Sender") %>%  # If Sender was in row names
  pivot_longer(-Sender, names_to = "Category", values_to = "Count")

# Plot
ggplot(long_filtered, aes(x = Sender, y = Count, fill = Category)) +
  geom_col(position = "stack") +
  labs(
    title = "Message Category Distribution by Sender (Excluding Neutral/Chit-chat)",
    x = "Sender",
    y = "Count of Messages",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```